Recruiting Chapter

### 2020-12-17 manuscript.Rmd
revising intro

---
title: "Recruiting Chapter"
author: "Ozan Jaquette"
bibliography: ./other/burd_recruiting_chapter.bib
output: bookdown::pdf_document2
csl: ./other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
---

@@

@@

@@

Each of these marketing/recruiting interventions -- and dozens of other interventions described in this book -- represent organizational behaviors that could be the subject of empirical research. However, the vast majority of research on enrollment management behaviors focuses on the final stages of the enrollment funnel, specifically which applicants are admitted and the use of financial aid "leveraging" to convert admits to enrollees [E.G.] <!-- [@RN3536; @RN3544; @RN3523; @RN4131; @RN2241; @RN3564; @RN1948] -->.  By contrast, the enrollment management industry expends substantial resources on marketing/recruiting activities that target earlier stages of the enrollment funnel [@RN4664] [GIVE SOME CONCRETE BUDGET NUMBERS INDICATING BALL PARK SPENDING?]. When we consider descriptions of enrollment management practice from investigative journalism [e.g., ], from enrollment management consulting firms and from this volume [e.g., ], it becomes apparent that -- with the notable exception of a small sociological case-study literature [e.g., ] -- the research community has ignores a great number of enrollment management practices that plausibly affect access to higher education. As a consequence, enrollment management remains an opaque industry to policymakers and the public. We argue that developing thoughtful policies about enrollment management depends on researchers collectively devloping empirical literatures that document enrollment management practices and evaluate the effects of these practices on opportunities for students.

We present this chapter as an example of empirical research about marketing/recruiting behavior. This chapter analyzes off-campus recruiting visits by a sample of public research universities and a sample of selective private colleges and universities in 2017. We collected visit data by "scraping" URLs (e.g., "Coming to a neighborhood near you") once per week from college/university admissions websites and -- for the sample of public research universities -- by issuring public records requests. Our analyses to date have focused on public research universities. As described in more detail below, 12 of the 15 [FIX NUMBERS, IT'S 17 PUBLICS] public research universities in our sample made far more out-of-state visits than in-state visits, a finding antithetical to the stated mission of most public research universities. These out-of-state visits focused on wealthy, predominantly white schools and communities. Additionally, most public research universities made a disproportionate number of out-of-state visits to private schools, but visited few in-state private schools.

Building on these findings, this chapter analyzes off-campus visits to private high schools by public research universities and by selective private colleges and universities. We utilize social network methods and emphasize comparisons within and between public and private postsecondary institutions. Analyses are informed by two broad, related questions. First, which types of private high schools receive visits from which types of colleges and universities? Second, to what do extent colleges and universities visit private high schools that share similar characteristics with the college (e.g., geographic region, religious affiliation, academic reputation, and racial composition)?

The motivation for comparing the behavior of public and private coleges/universities is based on the idea that (out-of-state) visits to private high schools are antithetical to the stated mission of public research universities. By contrast, recruiting visits to private high schools are consistent with the mission of private colleges and universities. From this perspective, out-of-state visits to private schools by public research universities is one example of "privatization." A weakness of the privatization literature is that few empirical studies compare the public and private institutions with respect to particular behaviors associated with privatization [@RN4543]. Therefore, by comparing the (mostly out-of-state) visits to private high schools by public research universities to those of private colleges and universities, we hope to develop new insights about the privatization of public higher education.

We use social network analysis (SNA) to analyze off-campus recruiting visits. The motivation for using SNA is that SNA privileges relationships between actorsQualitative scholarship on private high schools finds that college access is substantially determined by the quality of the relationship between a particular college and a particular private high school [@RN4321; @RN4407]. An off-campus recruiting visit is an indicator of a relationship between a college and a school that can be analyzed using standard tools of SNA.

[ADD ONE-PARAGRAPH OUTLINE?]

Our research analyzes off-campus recruiting visits by college and university admissions staff. So what do we know about off-campus recruiting? As is true for most aspects of enrollment management, much of what we know about off-campus recruiting comes from consulting firms (e.g., Ruffalo-Noel Levitz, EAB), professional associations (e.g., NACAC), and from practioner-oriented publications. Market research conceives of off-campus recruiting visits as a means of identifying prospects, deepening engagement with prospects already being targeted through mail/email, and maintaining relationships with guidance counselors at “feeder schools” [@RN4323; @RN4402; @RN4664]. With respect to expenditure, @RN4664 found that private (non-profit) 4-yr institutions spent an average of 17% of their undergraduate marketing/recruiting budget on "travel" to high schools and college fairs (as shown in Table X), a higher percentage of budget allocation than any other marketing/recruiting activity. Public institutions spent an average of 16% of their budget on travel, second only to "prospective student communications" at 17%. An emergent trend over the past decade -- partially a response to public universities seeking nonresident students -- has been the growth of "regional recruiters" who target specific metropolitan areas in the US and abroad. These regional recruiters may be college/university employees or they may be independent contractors who live in the metropolitan area they recruit.

@@

>>>>

---
title: "Recruiting Chapter"
author: "Ozan Jaquette"
bibliography: ./other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
---

@@

<!-- 
A pair of papers by Stevens provide interesting statements about the functions of higher education and the behavior of colleges and universities
Colleges and universities visit high schools that enroll desirable prospects, while high schools accept visits from colleges and universities that offer enrollment opportunities valued by some segment of the class. Drawing from a pair of papers by Mitchell Stevens [@RN2377; @RN4517], we argue that analysis of which high schools receive visits from which colleges and univeristies provides an opportunity how K-12 and postsecondary organizations form relationships that serve social reproduction.

This chapter presents a social network analysis of off-campus recruiting visits made by US colleges and universities to US private high schools. Colleges and universities visit high schools that enroll desirable prospects, while high schools accept visits from colleges and universities that offer enrollment opportunities valued by some segment of the class. An off-campus recruiting visit is an indicator of a relationship between a college and a school that can be analyzed using standard tools of social network analysis. We analyze visits made in 2017 from a sample of 25 selective private colleges and universities and a sample of 17 public research universities.  Analyses investigate which types of private high schools receive visits from which types of colleges and universities, emphasizing comparisons within and between public and private postsecondary institutions.

The motivation for these analyses is grounded within the sociology of higher education, particularly two theoretical and empirical syntheses developed by Mitchell Stevens and colleagues [@RN2377; @RN4517]. First, @RN2377 introduces metaphors that describe how sociology has conceived the functions of higher education. The stratification tradition conceives of higher education as a "sieve" -- an instrument that strains and separates -- that allocates people to occupational and social positions. This metaphor builds on "Max Weber's core insight that education has a dual character" [@RN2377, p. 129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to universities that confer privileged positions. The institutional theory tradition in sociology conceives of higher education as a "hub" that connects, pollinates, and legitimizes "prominent institutional sectors of modern societies" (e.g., professions, the philanthropic sector, corporate sector).  As a hub, higher education is "simultaneously a powerful and fragile social institution. On the one hand, higher education [is powerful because it] connects and reciprocally blesses various forms of privilege. Elite groups (e.g., high-status families, professions, politicians, scientists, intellectuals) and those aspiring to be a part of them use higher education to certify their legitimacy" [@RN2377, p. 137]. On the other hand, higher education is fragile because it depends on "patronage as a primary source of revenue."

@RN4517 introduces three alternative metaphors to describe the relationship between higher education and the American political economy. The "association" conceptualization, observed by @RN485, views colleges as voluntary (i.e., non-governmental) organizations founded by voluntary associations -- principally religious congregations and local business interests -- that rely "on the collective contributions and goodwill of their memberships" [@RN2377, p. 126]. Second, the "national service" configuration views higher education as a public-good engine of economic development and nation building and thus worthy of investment from federal and state governments. Third, "market" configuration conceives of higher education as a private good, suggesting that colleges should survive by competing for customers rather than rely on public funding. Each of these configurations is more or less associated with particular historical epochs and particular types of colleges and universities. For example, the association configuration defined US higher education through the 19th century, but remains salient for private colleges and universities.

<!-- 
The national service configuration is associated associated with massive public investment in higher education in the decades following World War II, while the market configuration is associated with declining state support for public universities since the 1980s.

These metaphors -- which describe higher education simultaneously serving social mobility and social reproduction and depending on patronage from the government, from voluntary associations, and private households -- are useful for comparing public flagship universities to selective private colleges and universities
-->

These metaphors are useful for comparing the comparing public flagship universities to selective private colleges and universities. Selective private colleges and universities depend on patronage from affluent households (market configuration) and philanthropic interests (associational configuration) and primarily serve social reproduction rather than social mobility. Historically, public flagship universities depended on government funding associated with the national service configuration and served social mobility rather than reproduction. Since the 1980s, reductions in public funding associated with the emergence of the market configuration encourage public flagship universities to prioritize patronage from affluent households and philanthropic interests. As suggested by the adage, "he who pays the piper calls the tune," reliance on affluent private-sector interests increases the likelihood that public flagship universities -- like selective private colleges and universities -- will primarily serve social reproduction rather than social mobility.
<!-- 
The extent to which higher education serves reproduction versus redistribution is sensitive to the external funding environment because universities depend on ``patronage as a primary revenue source'' [@RN2377, p. 138]. 
-->

We argue that the recruiting behavior of colleges and universities is an indicator of organizational enrollment priorities and, in turn, the extent to which enrollment priorities serve social mobility or social reproduction. However, access to concrete data on university recruiting behaviors has been a persistent barrier to empirical research. We collected data about off-campus recruiting visits made in 2017 by "scraping" URLs (e.g., "Coming to a neighborhood near you") once per week from college/university admissions websites and -- for a sample of public research universities -- by issuing public records requests. The first study from this data collection analyzed visits made by 15 public research universities [@RN4450]. The overwhelming majority of public research universities in our sample made more out-of-state visits than in-state visits, a finding antithetical to historic mission of public universities. These out-of-state visits focused on wealthy, predominantly white schools and communities, with a disproportionate number visits to out-of-state private high schools but relatively few visits to in-state private high schools.

Building on [@RN4450], this chapter analyzes off-campus recruiting visits to private high schools made in 2017 by a sample of 25 selective private colleges and universities and 17 public research universities. We utilize social network methods and emphasize comparisons within and between public and private postsecondary institutions. Analyses are informed by two broad, related questions. First, which types of private high schools receive visits from which types of colleges and universities? Second, to what do extent colleges and universities visit private high schools that share similar characteristics with the college (e.g., geographic region, religious affiliation, academic reputation, and racial composition)?

<!-- 
Our decision to compare public and private colleges/universities on the dimension of visits to private schools is motivated by the idea that recruiting from private high schools serves social reproduction rather than social mobility.

Our decision to compare public and private colleges/universities on the dimension of visits to private schools is motivated by the idea that recruiting from private high schools serves social reproduction rather than social mobility. Thus, visits to private schools are antithetical to the historic mission of public universities, especially given that the vast majority of visits are to out-of-state private schools. By contrast, recruiting from private high schools is consistent with the mission of selective private colleges and universities. From this perspective, visits to private schools by public research universities is one example of "privatization." A weakness of the empirical literature on privatization in higher education is that few studies public and private colleges/universities with respect to a behavior associated with privatization [@RN4543]. By comparing public and private colleges/universities on the behavioral dimension of recruiting visits to private schools, we hope to develop new insights about the privatization of public higher education.

Our decision to analyze visits to private schools is motivated by the idea that recruiting from private schools largely serves social reproduction [CITE A SOCIOLOGIST]. Thus, visits to private school are antithetical to the historic "social mobility mission" of public universities [@RN2269], particularly since most private school visits by public research universities are to out-of-state schools [@RN4450]. By contrast, recruiting from private high schools is consistent with the missions of selective private colleges and universities. From this perspective, private school visits by public research universities is an example of "privatization" that serves social reproduction.  A weakness of the empirical literature on privatization in higher education is that few studies public and private colleges/universities with respect to a behavior associated with privatization [@RN4543]. By comparing public and private colleges/universities on the behavioral dimension of recruiting visits to private schools, we hope to develop new insights about the privatization of public higher education.

Finally, although private school students are more likely than public school students to come from affluent, college-educated households, the population of US private schools contains tremendous variation in religious affiliation, academic reputation, racial and socioeconomic composition, amongst other dimensions [@RN4504]. Therefore, focusing solely on visits to private schools enables us to develop a nuanced analysis of which types of private high schools receive visits from which types colleges and universities. These nuances would be dificult to observe in an analysis that also included visits to public high schools.

<!-- 
We use social network analysis (SNA) to analyze off-campus recruiting visits. The motivation for using SNA is that SNA privileges relationships between actorsQualitative scholarship on private high schools finds that college access is substantially determined by the quality of the relationship between a particular college and a particular private high school [@RN4321; @RN4407]. An off-campus recruiting visit is an indicator of a relationship between a college and a school that can be analyzed using standard tools of SNA.
-->

@@

This section provides context for our analyses by, first, situating off-campus recruiting within the broader set of marketing and recruiting interventions. Second, we review what is known about off-campus recruiting from market research and from academic scholarship, with a particular focus on scholarship that examines the relationship between private high schools and colleges/universities.

@@

What do we know about off-campus recruiting from previous research? As is true for most aspects of enrollment management, much of what we know about off-campus recruiting comes from consulting firms (e.g., Ruffalo-Noel Levitz, EAB), professional associations (e.g., NACAC), and from practioner-oriented publications. Market research conceives of off-campus recruiting visits as a means of identifying prospects, deepening engagement with prospects already being targeted through mail/email, and maintaining relationships with guidance counselors at “feeder schools” [@RN4323; @RN4402; @RN4664]. With respect to expenditure, @RN4664 found that private (non-profit) 4-yr institutions spent an average of 17% of their undergraduate marketing/recruiting budget on "travel" to high schools and college fairs (as shown in Table X), a higher percentage of budget allocation than any other marketing/recruiting activity. Public institutions spent an average of 16% of their budget on travel, second only to "prospective student communications" at 17%. An emergent trend over the past decade -- partially a response to public universities seeking nonresident students -- has been the growth of "regional recruiters" who target specific metropolitan areas in the US and abroad. These regional recruiters may be college/university employees or they may be independent contractors who live in the metropolitan area they recruit.

@@

# Discussion

TEXT LA LA 
<!--

Each of these marketing/recruiting interventions -- and dozens of other interventions described in this book -- represent organizational behaviors that could be the subject of empirical research. However, the vast majority of research on enrollment management behaviors focuses on the final stages of the enrollment funnel, specifically which applicants are admitted and the use of financial aid "leveraging" to convert admits to enrollees [E.G.] [@RN3536; @RN3544; @RN3523; @RN4131; @RN2241; @RN3564; @RN1948]

.  By contrast, the enrollment management industry expends substantial resources on marketing/recruiting activities that target earlier stages of the enrollment funnel [@RN4664] [GIVE SOME CONCRETE BUDGET NUMBERS INDICATING BALL PARK SPENDING?]. When we consider descriptions of enrollment management practice from investigative journalism [e.g., ], from enrollment management consulting firms and from this volume [e.g., ], it becomes apparent that -- with the notable exception of a small sociological case-study literature [e.g., ] -- the research community has ignores a great number of enrollment management practices that plausibly affect access to higher education. As a consequence, enrollment management remains an opaque industry to policymakers and the public. We argue that developing thoughtful policies about enrollment management depends on researchers collectively devloping empirical literatures that document enrollment management practices and evaluate the effects of these practices on opportunities for students.
